### 2017-10-24 Draft2/Appending.Dynamic.tex
Rename variables

I will first implement appending for dynamic arrays. Let $\ListName$ be a dynamic array. The following definitions are used in the code:

\begin{description}
	\item[initial capacity] The capacity of a dynamic a array when one item is appended to it. I denote this $\InitCapacity$.\\
	(The capacity of any dynamic array with size zero should be zero.)
	\item[growth factor] The factor by which the current capacity is multiplied to get the new capacity when $\ListName$ grows. I denote this with $\GrowthFactor$.
\end{description}

\begin{algorithm}
	\begin{algorithmic}[1]
		\Procedure{$\FuncAppend$}{$L, item$}
			\If{$\ListName.\FieldFull$}
				\State $\ListName.\FuncGrow()$
			\EndIf

			\State $\ListName.\FieldHead[\ListName.\FieldHeadSize] \gets item$
			\State $\ListName.\FieldSize \gets \ListName.\FieldSize + 1$
		\EndProcedure
		\Statex
		\Procedure{$\FuncGrow$}{$L$}
			\If{$\ListName.\FieldEmpty$}
				\State $\ListName.\FieldBuffer \gets \FuncNewArray(\InitCapacity)$
			\Else
				\State $new\ buf \gets \FuncNewArray(\ListName.\FieldSize \times \GrowthFactor)$
				\State $\FuncArrayCopy(source: \ListName.\FieldBuffer, dest: new\ buf, len: \ListName.\FieldSize)$
				\State $\ListName.\FieldBuffer \gets new\ buf$
			\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

Before I analyze the time complexity of $\FuncAppend$, I consider a different method for measuring its cost. Suppose I start with an empty list and $n$ elements are appended. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $n$ appends, and denote it $\nwritesfn(n)$.

In the above code, one write is performed after the conditional, so it is apparent that $\nwritesfn(n) \geq n$ after $n$ appends. However, $\FuncGrow$ also does some writing, so in order to find a precise formula for $\nwritesfn(n)$, I need to analyze when $\FuncGrow$ is called. Consider the following lemma:

\begin{lemma}
\label{CapacitySequence}
	Let $\ListName$ by a dynamic array. As $n$ items are appended, the sequence of capacities $\ListName$ takes on is $$S_C = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$$.
\end{lemma}

\begin{proof}
	The capacity starts out at zero and becomes $\InitCapacity$ once one element is appended. From then on, it can only grow by a factor of $\GrowthFactor$ via $\FuncGrow$, so if $\GrowthFactor^i\InitCapacity$ is the current capacity then $\GrowthFactor^{i + 1}\InitCapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \GrowthFactor^i\InitCapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $n$, $k$ is the smallest integer for which $\GrowthFactor^k\InitCapacity \geq n$. Using this property, I now derive $k$.

	\begin{align*}
	\GrowthFactor^k\InitCapacity &\geq n\\
	\GrowthFactor^k &\geq \frac{n}{\InitCapacity}\\
	k &\geq \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity\\
	k &\geq \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity > k - 1\\
	k &= \big\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \big\rceil
	\end{align*}

	Then the final term in the sequence is $\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$, completing the proof.
\end{proof}

[make lemma]
Now, consider that to grow by $\GrowthFactor$, $\FuncGrow$ must be called. So $\GrowthFactor^k\InitCapacity$ being present in $S_C$ means $\FuncGrow$ was called at size $\GrowthFactor^{k - 1}\InitCapacity$ for $k \geq 1$. Then $\FuncGrow$ was called at the sizes $$S_R = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity$$

@@

\begin{align*}
0 + \InitCapacity + \GrowthFactor\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity &= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

Finally, counting the writes made for every item in $\FuncAppend$, I finally find an explicit formula for $\nwritesfn(n)$.

[ref1]
$$
\nwritesfn(n) = n + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

[ref2]
To make this expression easier to work with, I will write it in big-P notation to approximate it for large $n$. First I note that $\log_{\GrowthFactor} k \leq \lceil \log_{\GrowthFactor} k \rceil < \left( \log_{\GrowthFactor} k \right) + 1$ for any $k > 0$, so $k \leq \GrowthFactor^{\lceil \log_{\GrowthFactor} k \rceil} < \GrowthFactor{k}$. Then

\begin{align*}
\nwritesfn(n) &= n + \left( \frac{\GrowthFactor^{\big\lceil \log_{\GrowthFactor} \left( \frac {n} {\InitCapacity} \right) \big\rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
n + \left( \frac{\frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \nwritesfn(n) &< n + \left( \frac{\GrowthFactor\frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
n + \left( \frac{n - \InitCapacity}{\GrowthFactor - 1} \right) \leq \nwritesfn(n) &< n + \left( \frac{\GrowthFactorn - \InitCapacity}{\GrowthFactor - 1} \right)\\
P\left( n + \left( \frac{n - \InitCapacity}{\GrowthFactor - 1} \right) \right) \leq P\left( \nwritesfn(n) \right) &< P\left( n + \left( \frac{\GrowthFactorn - \InitCapacity}{\GrowthFactor - 1} \right) \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) n \right) \leq P\left( \nwritesfn(n) \right) &< P\left( \left( \frac{2\GrowthFactor - 1}{\GrowthFactor - 1} \right) n \right)
\end{align*}

@@

I wish to find the space allocated while appending $n$ items to a dynamic array. I denote this quantity $\spacefn(n)$, and define it as the total length of all arrays allocated during $n$ $\FuncAppend$ calls. I now derive a formula for $\spacefn(n)$.

From \ref{CapacitySequence} I know the exact capacities $\ListName$ takes on as $n$ items are appended. Dynamic arrays also have the property that $i \in S_C$ iff an array of length $i$ was allocated. Then the total length of arrays allocated is

\begin{align*}
\spacefn(n) &= \sum_{i \in S_C} i\\
&= 0 + \InitCapacity + \GrowthFactor\InitCapacity + \GrowthFactor^2\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity\\
&= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil + 1} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

@@

\begin{align*}
\left( \frac{\GrowthFactor \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \spacefn(n) &< \left( \frac{\GrowthFactor^2 \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
P\left( \left( \frac{\GrowthFactor \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right) \leq P\left( \spacefn(n) \right) &< P\left( \left( \frac{\GrowthFactor^2 \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) n \right) \leq P\left( \spacefn(n) \right) &< P\left( \left( \frac{\GrowthFactor^2}{\GrowthFactor - 1} \right) n \right)
\end{align*}

>>>>

I will first implement appending for dynamic arrays. Let $\VarList$ be a dynamic array. The following definitions are used in the code:

\begin{description}
	\item[initial capacity] The capacity of a dynamic a array when one item is appended to it. I denote this $\VarInitCapacity$.\\
	(The capacity of any dynamic array with size zero should be zero.)
	\item[growth factor] The factor by which the current capacity is multiplied to get the new capacity when $\VarList$ grows. I denote this with $\VarGrowthFactor$.
\end{description}

\begin{algorithm}
	\begin{algorithmic}[1]
		\Procedure{$\FuncAppend$}{$L, item$}
			\If{$\VarList.\FieldFull$}
				\State $\VarList.\FuncGrow()$
			\EndIf

			\State $\VarList.\FieldHead[\VarList.\FieldHeadSize] \gets item$
			\State $\VarList.\FieldSize \gets \VarList.\FieldSize + 1$
		\EndProcedure
		\Statex
		\Procedure{$\FuncGrow$}{$L$}
			\If{$\VarList.\FieldEmpty$}
				\State $\VarList.\FieldBuffer \gets \FuncNewArray(\VarInitCapacity)$
			\Else
				\State $new\ buf \gets \FuncNewArray(\VarList.\FieldSize \times \VarGrowthFactor)$
				\State $\FuncArrayCopy(source: \VarList.\FieldBuffer, dest: new\ buf, len: \VarList.\FieldSize)$
				\State $\VarList.\FieldBuffer \gets new\ buf$
			\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

Before I analyze the time complexity of $\FuncAppend$, I consider a different method for measuring its cost. Suppose I start with an empty list and $n$ elements are appended. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $n$ appends, and denote it $\FuncWrites(n)$.

In the above code, one write is performed after the conditional, so it is apparent that $\FuncWrites(n) \geq n$ after $n$ appends. However, $\FuncGrow$ also does some writing, so in order to find a precise formula for $\FuncWrites(n)$, I need to analyze when $\FuncGrow$ is called. Consider the following lemma:

\begin{lemma}
\label{CapacitySequence}
	Let $\VarList$ by a dynamic array. As $n$ items are appended, the sequence of capacities $\VarList$ takes on is $$S_C = 0,\ \VarInitCapacity,\ \VarGrowthFactor\VarInitCapacity,\ \VarGrowthFactor^2\VarInitCapacity,\ \ldots\ \VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil}\VarInitCapacity$$.
\end{lemma}

\begin{proof}
	The capacity starts out at zero and becomes $\VarInitCapacity$ once one element is appended. From then on, it can only grow by a factor of $\VarGrowthFactor$ via $\FuncGrow$, so if $\VarGrowthFactor^i\VarInitCapacity$ is the current capacity then $\VarGrowthFactor^{i + 1}\VarInitCapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \VarGrowthFactor^i\VarInitCapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $n$, $k$ is the smallest integer for which $\VarGrowthFactor^k\VarInitCapacity \geq n$. Using this property, I now derive $k$.

	\begin{align*}
	\VarGrowthFactor^k\VarInitCapacity &\geq n\\
	\VarGrowthFactor^k &\geq \frac{n}{\VarInitCapacity}\\
	k &\geq \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity\\
	k &\geq \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity > k - 1\\
	k &= \big\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \big\rceil
	\end{align*}

	Then the final term in the sequence is $\VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil}\VarInitCapacity$, completing the proof.
\end{proof}

[make lemma]
Now, consider that to grow by $\VarGrowthFactor$, $\FuncGrow$ must be called. So $\VarGrowthFactor^k\VarInitCapacity$ being present in $S_C$ means $\FuncGrow$ was called at size $\VarGrowthFactor^{k - 1}\VarInitCapacity$ for $k \geq 1$. Then $\FuncGrow$ was called at the sizes $$S_R = 0,\ \VarInitCapacity,\ \VarGrowthFactor\VarInitCapacity,\ \VarGrowthFactor^2\VarInitCapacity,\ \ldots\ \VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil - 1}\VarInitCapacity$$

@@

\begin{align*}
0 + \VarInitCapacity + \VarGrowthFactor\VarInitCapacity + \ldots + \VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil - 1}\VarInitCapacity &= \left( \frac{\VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}

Finally, counting the writes made for every item in $\FuncAppend$, I finally find an explicit formula for $\FuncWrites(n)$.

[ref1]
$$
\FuncWrites(n) = n + \left( \frac{\VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
$$

[ref2]
To make this expression easier to work with, I will write it in big-P notation to approximate it for large $n$. First I note that $\log_{\VarGrowthFactor} k \leq \lceil \log_{\VarGrowthFactor} k \rceil < \left( \log_{\VarGrowthFactor} k \right) + 1$ for any $k > 0$, so $k \leq \VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} k \rceil} < \VarGrowthFactor{k}$. Then

\begin{align*}
\FuncWrites(n) &= n + \left( \frac{\VarGrowthFactor^{\big\lceil \log_{\VarGrowthFactor} \left( \frac {n} {\VarInitCapacity} \right) \big\rceil} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity\\
n + \left( \frac{\frac{n}{\VarInitCapacity} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \leq \FuncWrites(n) &< n + \left( \frac{\VarGrowthFactor\frac{n}{\VarInitCapacity} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity\\
n + \left( \frac{n - \VarInitCapacity}{\VarGrowthFactor - 1} \right) \leq \FuncWrites(n) &< n + \left( \frac{\VarGrowthFactor{n} - \VarInitCapacity}{\VarGrowthFactor - 1} \right)\\
P\left( n + \left( \frac{n - \VarInitCapacity}{\VarGrowthFactor - 1} \right) \right) \leq P\left( \FuncWrites(n) \right) &< P\left( n + \left( \frac{\VarGrowthFactor{n} - \VarInitCapacity}{\VarGrowthFactor - 1} \right) \right)\\
P\left( \left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \right) \leq P\left( \FuncWrites(n) \right) &< P\left( \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n \right)
\end{align*}

@@

I wish to find the space allocated while appending $n$ items to a dynamic array. I denote this quantity $\FuncSpace(n)$, and define it as the total length of all arrays allocated during $n$ $\FuncAppend$ calls. I now derive a formula for $\FuncSpace(n)$.

From \ref{CapacitySequence} I know the exact capacities $\VarList$ takes on as $n$ items are appended. Dynamic arrays also have the property that $i \in S_C$ iff an array of length $i$ was allocated. Then the total length of arrays allocated is

\begin{align*}
\FuncSpace(n) &= \sum_{i \in S_C} i\\
&= 0 + \VarInitCapacity + \VarGrowthFactor\VarInitCapacity + \VarGrowthFactor^2\VarInitCapacity + \ldots + \VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil}\VarInitCapacity\\
&= \left( \frac{\VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} n - \log_{\VarGrowthFactor} \VarInitCapacity \rceil + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}

@@

\begin{align*}
\left( \frac{\VarGrowthFactor \frac{n}{\VarInitCapacity} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \leq \FuncSpace(n) &< \left( \frac{\VarGrowthFactor^2 \frac{n}{\VarInitCapacity} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity\\
P\left( \left( \frac{\VarGrowthFactor \frac{n}{\VarInitCapacity} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \right) \leq P\left( \FuncSpace(n) \right) &< P\left( \left( \frac{\VarGrowthFactor^2 \frac{n}{\VarInitCapacity} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \right)\\
P\left( \left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \right) \leq P\left( \FuncSpace(n) \right) &< P\left( \left( \frac{\VarGrowthFactor^2}{\VarGrowthFactor - 1} \right) n \right)
\end{align*}

### 2017-10-24 Draft2/Appending.Growth.tex
Rename variables

I again start off by finding the write cost when $n$ are appended. I give the write cost function a slightly different name, $\nwritesnewfn(n)$, so I can compare it to $\nwritesfn(n)$ later. I also use $\GrowthFactornew$ and $\InitCapacitynew$ instead of $\GrowthFactor$ and $\InitCapacity$ for the growth factor and initial capacity, respectively.

@@

\begin{lemma}
	The capacity of a growth array grows by a constant factor $\GrowthFactornew$ after the first time it is called.
\end{lemma}

@@

	For the base case, consider the second time $\FuncGrow$ is called: the current capacity is $\InitCapacitynew$, and the next (total) capacity should be $\GrowthFactornew\InitCapacitynew$. Then a buffer of size $(\GrowthFactornew - 1)\InitCapacitynew$ should be appended, and that is what the algorithm does.

	For $i \geq 2$, given the algorithm behaves correctly the $i$th time it is called, I wish to show it behaves correctly the $(i + 1)$th time. From the inductive hypothesis, when the algorithm is called the $(i + 1)$th time, the current capacity is $\GrowthFactornew^i\InitCapacitynew$ and the head capacity is $(\GrowthFactornew^i - \GrowthFactornew^{i - 1})\InitCapacitynew$. The next (total) capacity should be $\GrowthFactornew^{i + 1}\InitCapacitynew$, and the next head capacity should be $(\GrowthFactornew^{i + 1} - \GrowthFactornew^i)\InitCapacitynew$, which is $\GrowthFactornew$ times the current head capacity. The algorithm determines the capacity of the next buffer in this manner, proving the inductive step and completing the induction.
\end{proof}

@@

Let $\nwritesgrowfn(n)$ denote the total number of writes made by $\FuncGrow$. From [make lemma], I know that $\FuncGrow$ is called at the sizes

$$
S_R = 0,\ \InitCapacitynew,\ \GrowthFactornew\InitCapacitynew,\ \GrowthFactornew^2\InitCapacitynew,\ \ldots\ \GrowthFactornew^{\lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil - 1}\InitCapacitynew
$$

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\nwritesfn(k)$. Since $\FuncGrow$ is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Then the formula for $\nwritesgrowfn(n)$ is

$$
\nwritesgrowfn(n) = k + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} k - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

where $\InitCapacity$ $\GrowthFactor$ are the capacity of the \textit{tail} (not the growth array). I finally conclude that

$$
\nwritesnewfn(n) = n + k + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} k - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

@@

\begin{align*}
k &= \lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil\\
O(k) &= O(\log n)
\end{align*}

Then, it is shown that the $k$-term disappears when analyzing the big-P complexity of $\nwritesnewfn(n)$, leaving just $P(n)$:

\begin{align*}
P(\nwritesnewfn(n)) &= P(n + \nwritesfn(k))\\
&= P(n) + P(\nwritesfn(k))\\
&= P(n) + O(\log n)\\
&= P(n)
\end{align*}

@@

% Should this be C(n) instead of C?
The space function for $n$ $\FuncAppend$ calls on a growth array is denoted $\spacenewfn(n)$. Since growth arrays never throw away item buffers, if the current capacity is $C$ then the total length of item buffers is also $C$. $\spacenewfn(n)$ is usually greater than $C$, however; this is because growth arrays not only allocate item buffers, but stores references to them in the tail. Thus the space the tail allocates must also be determined.

I determined in [...] that if $n$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Since the tail is a dynamic array, it allocates $\spacefn(k)$ space, where $\spacefn$ is the space function of the tail. Through a similar method to [...] it can be shown $O(\spacefn(k)) = O(k) = O(\log n)$. Then % TODO: P-analysis should come after the explicit formula.

\begin{align*}
\spacenewfn(n) = P(2^{\lceil \log_2 n \rceil}) = O(n)
\end{align*}

>>>>

I again start off by finding the write cost when $n$ are appended. I give the write cost function a slightly different name, $\FuncWritesNew(n)$, so I can compare it to $\FuncWrites(n)$ later. I also use $\VarGrowthFactorNew$ and $\VarInitCapacityNew$ instead of $\VarGrowthFactor$ and $\VarInitCapacity$ for the growth factor and initial capacity, respectively.

@@

\begin{lemma}
	The capacity of a growth array grows by a constant factor $\VarGrowthFactorNew$ after the first time it is called.
\end{lemma}

@@

	For the base case, consider the second time $\FuncGrow$ is called: the current capacity is $\VarInitCapacityNew$, and the next (total) capacity should be $\VarGrowthFactorNew\VarInitCapacityNew$. Then a buffer of size $(\VarGrowthFactorNew - 1)\VarInitCapacityNew$ should be appended, and that is what the algorithm does.

	For $i \geq 2$, given the algorithm behaves correctly the $i$th time it is called, I wish to show it behaves correctly the $(i + 1)$th time. From the inductive hypothesis, when the algorithm is called the $(i + 1)$th time, the current capacity is $\VarGrowthFactorNew^i\VarInitCapacityNew$ and the head capacity is $(\VarGrowthFactorNew^i - \VarGrowthFactorNew^{i - 1})\VarInitCapacityNew$. The next (total) capacity should be $\VarGrowthFactorNew^{i + 1}\VarInitCapacityNew$, and the next head capacity should be $(\VarGrowthFactorNew^{i + 1} - \VarGrowthFactorNew^i)\VarInitCapacityNew$, which is $\VarGrowthFactorNew$ times the current head capacity. The algorithm determines the capacity of the next buffer in this manner, proving the inductive step and completing the induction.
\end{proof}

@@

Let $\FuncWritesByGrow(n)$ denote the total number of writes made by $\FuncGrow$. From [make lemma], I know that $\FuncGrow$ is called at the sizes

$$
S_R = 0,\ \VarInitCapacityNew,\ \VarGrowthFactorNew\VarInitCapacityNew,\ \VarGrowthFactorNew^2\VarInitCapacityNew,\ \ldots\ \VarGrowthFactorNew^{\lceil \log_{\VarGrowthFactorNew} n - \log_{\VarGrowthFactorNew} \VarInitCapacityNew \rceil - 1}\VarInitCapacityNew
$$

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\FuncWrites(k)$. Since $\FuncGrow$ is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\VarGrowthFactorNew} n - \log_{\VarGrowthFactorNew} \VarInitCapacityNew \rceil$. Then the formula for $\FuncWritesByGrow(n)$ is

$$
\FuncWritesByGrow(n) = k + \left( \frac{\VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} k - \log_{\VarGrowthFactor} \VarInitCapacity \rceil} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
$$

where $\VarInitCapacity$ $\VarGrowthFactor$ are the capacity of the \textit{tail} (not the growth array). I finally conclude that

$$
\FuncWritesNew(n) = n + k + \left( \frac{\VarGrowthFactor^{\lceil \log_{\VarGrowthFactor} k - \log_{\VarGrowthFactor} \VarInitCapacity \rceil} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
$$

@@

\begin{align*}
k &= \lceil \log_{\VarGrowthFactorNew} n - \log_{\VarGrowthFactorNew} \VarInitCapacityNew \rceil\\
O(k) &= O(\log n)
\end{align*}

Then, it is shown that the $k$-term disappears when analyzing the big-P complexity of $\FuncWritesNew(n)$, leaving just $P(n)$:

\begin{align*}
P(\FuncWritesNew(n)) &= P(n + \FuncWrites(k))\\
&= P(n) + P(\FuncWrites(k))\\
&= P(n) + O(\log n)\\
&= P(n)
\end{align*}

@@

% Should this be C(n) instead of C?
The space function for $n$ $\FuncAppend$ calls on a growth array is denoted $\FuncSpaceNew(n)$. Since growth arrays never throw away item buffers, if the current capacity is $C$ then the total length of item buffers is also $C$. $\FuncSpaceNew(n)$ is usually greater than $C$, however; this is because growth arrays not only allocate item buffers, but stores references to them in the tail. Thus the space the tail allocates must also be determined.

I determined in [...] that if $n$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\VarGrowthFactorNew} n - \log_{\VarGrowthFactorNew} \VarInitCapacityNew \rceil$. Since the tail is a dynamic array, it allocates $\FuncSpace(k)$ space, where $\FuncSpace$ is the space function of the tail. Through a similar method to [...] it can be shown $O(\FuncSpace(k)) = O(k) = O(\log n)$. Then % TODO: P-analysis should come after the explicit formula.

\begin{align*}
\FuncSpaceNew(n) = P(2^{\lceil \log_2 n \rceil}) = O(n)
\end{align*}

### 2017-10-24 Draft2/FieldsProperties.tex
Rename variables

If $\ListName$ is a dynamic array, then it is assumed to have the following fields:

\begin{itemize}
	\item $\ListName.\FieldBuffer$ - The \textbf{buffer}, or raw array, that $\ListName$ stores its items in.
	\item $\ListName.\FieldSize$ - The number of items in $\ListName$.
\end{itemize}

As a dynamic array, $\ListName$ is also given the following properties. $:=$ denotes a definition, as opposed to $=$ which checks equivalence. Functions that return boolean values are suffixed with ?.

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns the capacity of $\ListName$}
		\State $\ListName.\FieldCapacity := \ListName.\FieldBuffer.\FieldLength$
		\State
		\LineComment{Returns whether $\ListName$ is full}
		\State $\ListName.\FieldFull := \ListName.\FieldSize = \ListName.\FieldCapacity$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
	\begin{algorithmic}
		\State $\ListName.\FieldBuffer \gets \FuncNewArray(0)$
		\State $\ListName.\FieldSize \gets 0$
	\end{algorithmic}
\end{algorithm}

If $\ListName$ is a growth array, then it is assumed to have the following fields:

\begin{itemize}
	\item $\ListName.\FieldHead$ - The \textbf{head} of $\ListName$. It returns the buffer we are currently adding items to.
	\item $\ListName.\FieldTail$ - The \textbf{tail} of $\ListName$. It returns a dynamic array of buffers that have already been filled. This can be thought of as a two-dimensional array.
	\item $\ListName.\FieldSize$ - The number of items in $\ListName$.
	\item $\ListName.\FieldCapacity$ - The \textbf{capacity} of $\ListName$. It returns the maximum number of items $\ListName$ can hold without resizing.
\end{itemize}

As a growth array, $\ListName$ is also given the following properties:

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns whether $\ListName$ is empty}
		\State $\ListName.\FieldEmpty := \ListName.\FieldSize = 0$
		\State
		\LineComment{Returns whether $\ListName$ is full}
		\State $\ListName.\FieldFull := \ListName.\FieldSize = \ListName.\FieldCapacity$
		\State
		\LineComment{Returns the capacity of $\FieldHead$}
		\State $\ListName.\FieldHeadCapacity := \ListName.\FieldHead.\FieldLength$
		\State
		\LineComment{Returns the size of $\FieldHead$}
		\LineComment{\textbf{Rationale:} $\FieldCapacity - \FieldHeadCapacity$ is the total capacity of the buffers in $\FieldTail$.}
		\LineComment{Then, $\FieldSize - \left( \FieldCapacity - \FieldHeadCapacity \right)$ is the number of items that were added}
		\LineComment{after depleting the buffers in $\FieldTail$.}
		\State $\ListName.\FieldHeadSize := \ListName.\FieldSize - \left(\ListName.\FieldCapacity - \ListName.\FieldHeadCapacity\right)$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
	\begin{algorithmic}
		\State $\ListName.\FieldHead \gets \FuncNewArray(0)$
		\State $\ListName.\FieldTail \gets \FuncNewDynamicArray()$
		\State $\ListName.\FieldSize \gets 0$
		\State $\ListName.\FieldCapacity \gets 0$
	\end{algorithmic}
\end{algorithm}

>>>>

If $\VarList$ is a dynamic array, then it is assumed to have the following fields:

\begin{itemize}
	\item $\VarList.\FieldBuffer$ - The \textbf{buffer}, or raw array, that $\VarList$ stores its items in.
	\item $\VarList.\FieldSize$ - The number of items in $\VarList$.
\end{itemize}

As a dynamic array, $\VarList$ is also given the following properties. $:=$ denotes a definition, as opposed to $=$ which checks equivalence. Functions that return boolean values are suffixed with ?.

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns the capacity of $\VarList$}
		\State $\VarList.\FieldCapacity := \VarList.\FieldBuffer.\FieldLength$
		\State
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
	\begin{algorithmic}
		\State $\VarList.\FieldBuffer \gets \FuncNewArray(0)$
		\State $\VarList.\FieldSize \gets 0$
	\end{algorithmic}
\end{algorithm}

If $\VarList$ is a growth array, then it is assumed to have the following fields:

\begin{itemize}
	\item $\VarList.\FieldHead$ - The \textbf{head} of $\VarList$. It returns the buffer we are currently adding items to.
	\item $\VarList.\FieldTail$ - The \textbf{tail} of $\VarList$. It returns a dynamic array of buffers that have already been filled. This can be thought of as a two-dimensional array.
	\item $\VarList.\FieldSize$ - The number of items in $\VarList$.
	\item $\VarList.\FieldCapacity$ - The \textbf{capacity} of $\VarList$. It returns the maximum number of items $\VarList$ can hold without resizing.
\end{itemize}

As a growth array, $\VarList$ is also given the following properties:

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns whether $\VarList$ is empty}
		\State $\VarList.\FieldEmpty := \VarList.\FieldSize = 0$
		\State
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
		\State
		\LineComment{Returns the capacity of $\FieldHead$}
		\State $\VarList.\FieldHeadCapacity := \VarList.\FieldHead.\FieldLength$
		\State
		\LineComment{Returns the size of $\FieldHead$}
		\LineComment{\textbf{Rationale:} $\FieldCapacity - \FieldHeadCapacity$ is the total capacity of the buffers in $\FieldTail$.}
		\LineComment{Then, $\FieldSize - \left( \FieldCapacity - \FieldHeadCapacity \right)$ is the number of items that were added}
		\LineComment{after depleting the buffers in $\FieldTail$.}
		\State $\VarList.\FieldHeadSize := \VarList.\FieldSize - \left(\VarList.\FieldCapacity - \VarList.\FieldHeadCapacity\right)$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
	\begin{algorithmic}
		\State $\VarList.\FieldHead \gets \FuncNewArray(0)$
		\State $\VarList.\FieldTail \gets \FuncNewDynamicArray()$
		\State $\VarList.\FieldSize \gets 0$
		\State $\VarList.\FieldCapacity \gets 0$
	\end{algorithmic}
\end{algorithm}
